---
title : ARIMAX and Granger Causality
output:
  html_document:
    toc: yes
  html_notebook:
    fig_caption : yes
    highlight   : tango
    theme       : united
    toc         : yes
    toc_float   : yes
---

# Problem Description

Predict weekly sales quantity of the item, using retail store data

## Reading and Understanding Data

### Clear the Global Environment
```{r}
rm(list=ls(all=TRUE))
```

### Load required R library
```{r}
library(caret)
library(forecast)
library(DMwR)
```

###  Read ArimaxData.csv file
```{r,}
PATH = getwd()
setwd(PATH)
data = read.csv("ArimaxData.csv")
rm(PATH)
```

### Look at the first and last 6 records using head and tail commands
```{r, echo=T}
head(data, 6)

tail(data, 6)
```

### Explore and understand the data

```{r, echo=T}
## Dimension of the Data set
dim(data)

## Look at the summary statistics
summary(data)
```

### Look at the structure of the data
```{r, echo=T}
str(data)
```

### Convert attributes into appropriate type
```{r, echo=T}
data$temperaturebins = as.factor(data$temperaturebins)
str(data)
```

```{r, echo=T}
# Summary of the data
summary(data)

# Re-look at the first 6 records
head(data)
```

### ARIMAX accepts only numerics attribute. So, convert categorial attributes to numeric using dummification
* Since the dataframe   both Categorical and Numeric datatypes, Create two arrays having Numeric and Categorical Column names.
* Dummify categorical columns
* Merge Numeric columns with dummified columns.
```{r, echo=T}

names(data)
c("weeklyPromotion", "temperaturebins")
setdiff(names(data), c("weeklyPromotion", "temperaturebins"))

dataNum = data[,setdiff(names(data), c("weeklyPromotion", "temperaturebins"))]
head(dataNum)


dummyModel  = dummyVars(~weeklyPromotion+temperaturebins, data)
dataCat2Num = predict(dummyModel, data)

head(dataCat2Num)

data = cbind(dataNum, dataCat2Num)
head(data)
str(data)

```

### Split the data into Train and Test.
```{r, echo=T}
noOfWeekstoForcast = 4
trainRowID = 1:(nrow(data) - noOfWeekstoForcast)
trainData = data[trainRowID,]
testData = data[-trainRowID,]
rm(trainRowID, data)
```

### Verify Train and Test counts
```{r, echo=T}
nrow(trainData)
nrow(testData)
```

### Create Timeseries Objects
```{r, echo=T}
trainTS = ts(trainData$quantityPerWeek, frequency = 53, start = c(2014, 1))
testTS = ts(testData$quantityPerWeek, frequency = 53, start = c(2016, 46))
```

### Plot train Data
```{r, echo=T}
ts.plot(trainTS)
plot(decompose(trainTS))
```

# Model Building

## Holt-Winters Model

### Build Holt-Winters Model
```{r, echo=T}
hw_Model = HoltWinters(trainTS)
hw_Model
```

### Predictions of Hold-Winters on train data
```{r, echo=T}
train_HW_pred = hw_Model$fitted
train_HW_pred
```
### Performance of Hold-Winters on train data
```{r, echo=T}
err_HW_Train = regr.eval(trainData[nrow(trainData)-53:nrow(trainData),]$quantityPerWeek, train_HW_pred[,1])
err_HW_Train
```

### Forcasting using the model
```{r, echo=T}
test_HW_pred = forecast(hw_Model, noOfWeekstoForcast)
test_HW_pred
plot(test_HW_pred)

```

### Performance on Test data
```{r, echo=T}
err_HW_Test = regr.eval(testData$quantityPerWeek, data.frame(test_HW_pred)$Point.Forecast)
err_HW_Test
```


## ARIMA model

### Build Arima Model using auto.arima()
```{r, echo=T}
arima_Model = auto.arima(trainTS)
summary(arima_Model)
arimaorder(arima_Model)
```

### Predictions on Train data
```{r, echo=T}
train_Arima_Pred = fitted(arima_Model)
train_Arima_Pred
```

### Performance on Train data
```{r, echo=T}
err_Arima_Train = regr.eval(trainData$quantityPerWeek, data.frame(train_Arima_Pred)$Point.Forecast)
err_Arima_Train
```

### Forecast using the Arima model
```{r, echo=T}
test_Arima_Pred = forecast(arima_Model, h=noOfWeekstoForcast)
test_Arima_Pred

plot(test_Arima_Pred)
```

### Performance on Train data
```{r, echo=T}
err_Arima_Test = regr.eval(testData$quantityPerWeek, data.frame(test_Arima_Pred)$Point.Forecast)
err_Arima_Test

```

## Arimax Model

### Build a Arimax model using exogenous variables
```{r, echo=T}
arimax_Model_1 = auto.arima(trainTS, xreg=as.matrix(trainData[,c(4,5,8,13,15,16)]))
summary(arimax_Model_1)
arimaorder(arimax_Model_1)
```


### Predictions on Train data
```{r, echo=T}
train_Arimax_Pred_1 = fitted(arimax_Model_1)
train_Arimax_Pred_1
```


### Performance on Train data
```{r, echo=T}
err_Arimax_Train_1 = regr.eval(trainData$quantityPerWeek, train_Arimax_Pred_1)
err_Arimax_Train_1
```


### Forecast and plot using Arimax model
```{r, echo=T}
colnames(testData)
test_Arimax_Pred_1 = forecast(arimax_Model_1, h=noOfWeekstoForcast, xreg=as.matrix(testData[,c(4,5,8,13,15,16)]))
test_Arimax_Pred_1

plot(test_Arimax_Pred_1)
```


### Performance on Test data
```{r, echo=T}
err_Arimax_Test_1 = regr.eval(testData$quantityPerWeek, data.frame(test_Arimax_Pred_1)$Point.Forecast)
err_Arimax_Test_1

```


### Performance on all models on Train and Test Data
```{r, echo=T}
#H oltwinters

err_HW_Train
err_HW_Test

# Arima
err_Arima_Train
err_Arima_Test

#Arimax
err_Arimax_Train_1
err_Arimax_Test_1
```

## Improving the ARIMAX model

### import lmtest and vars libraries 
```{r, echo=T}
library(lmtest)

#install.packages("vars")
library(vars)

ts.plot(as.ts(trainData$NoOfWeeklyPromotion), as.ts(trainData$quantityPerWeek))
```

### plot the other numeric variables
```{r, echo=T}
ts.plot(as.ts(scale(trainData$NoOfWeeklyPromotion) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(trainData$FractionOfWeeklyPriceChange) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(trainData$AverageWeeklyPriceChange) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(trainData$AverageWeeklyPriceChangePer) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(trainData$meanTemperatureOfWeek) + 5), as.ts(scale(trainData$quantityPerWeek)))
```


### Number of differences required for target variable to be stationary
```{r, echo=T}
ndiffs(trainData$quantityPerWeek)
```
```{r, echo=T}
ndiffs(trainData$NoOfWeeklyPromotion)
ndiffs(trainData$FractionOfWeeklyPriceChange)
ndiffs(trainData$AverageWeeklyPriceChange)
ndiffs(trainData$AverageWeeklyPriceChangePer)
ndiffs(trainData$meanTemperatureOfWeek)
```

```{r, echo=T}
ts.plot(as.ts(scale(diff(trainData$FractionOfWeeklyPriceChange)) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(diff(trainData$AverageWeeklyPriceChange)) + 5), as.ts(scale(trainData$quantityPerWeek)))
ts.plot(as.ts(scale(diff(trainData$AverageWeeklyPriceChangePer)) + 5), as.ts(scale(trainData$quantityPerWeek)))
```
```{r, echo=T}
VARselect(trainData$quantityPerWeek, exogen=trainData[,c(4,5,6,7,8)])
```

```{r, echo=T}
grangertest(quantityPerWeek ~ NoOfWeeklyPromotion,data=trainData, order=10)
grangertest(trainData[2:151,]$quantityPerWeek ~ diff(FractionOfWeeklyPriceChange, lag=1),data=trainData, order=10)
grangertest(trainData[2:151,]$quantityPerWeek ~ diff(AverageWeeklyPriceChange,lag=1),data=trainData, order=10)
grangertest(trainData[2:151,]$quantityPerWeek ~ diff(AverageWeeklyPriceChangePer,lag=1),data=trainData, order=10)
grangertest(quantityPerWeek ~ meanTemperatureOfWeek,data=trainData, order=10)
```


```{r, echo=T}
arimax_Model_2 = auto.arima(trainTS, xreg=as.matrix(trainData[,c(5,8,6)]))

summary(arimax_Model_2)
arimaorder(arimax_Model_2)
```

```{r, echo=T}
train_Arimax_Pred_2 = fitted(arimax_Model_2)
train_Arimax_Pred_2

err_Arimax_Train_2 = regr.eval(trainData$quantityPerWeek, train_Arimax_Pred_2)
err_Arimax_Train_2
```

```{r, echo=T}

test_Arimax_Pred_2 = forecast(arimax_Model_2, h=noOfWeekstoForcast, xreg=as.matrix(testData[,c(5,8,6)]))
test_Arimax_Pred_2

plot(test_Arimax_Pred_2)
```


```{r, echo=T}
err_Arimax_Test_2 = regr.eval(testData$quantityPerWeek, data.frame(test_Arimax_Pred_2)$Point.Forecast)
err_Arimax_Test_2
```


```{r, echo=T}
# Performance of all models on both train and test
err_HW_Train
err_HW_Test

err_Arima_Train
err_Arima_Test

err_Arimax_Train_1
err_Arimax_Test_1

#Arimax
err_Arimax_Train_2
err_Arimax_Test_2
```



