---
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}
rm(list = ls(all=TRUE))
```

# Problem
* To predict the Employment rate in Wisconsin.

# Data
* Monthly Milk Production in US (Pounds per cow)
* Employment rate in Wisconsin (In Thousands)

# Problem Understanding and Approach
* As employment rate in wisconsin is numeric attribute it is regression problem.
* As there is only one independent attribute, lets apply simple linear regression.

# Reading the Data
```{r}
data = read.csv(file="srdata.csv", header=TRUE)
```

# Look at first 6 and last 6 records
# Summary statistics
```{r}
head(data)
tail(data)

summary(data)
```

# Simple Linear Regression
* Ignore Year_Month attribute. 
* Building a Simple Linear Regression model to with Employment data as target and Milk Production as independent
```{r}
lr = lm(Wisconsin_Employment_Rate~Milk_Production, data)
summary(lr)
```

# Understanding the model
* We can observe that R^2^ is decent for this problem.
* p-value is low for Milk_Production attribute.

# Observations

Something look fishy, because target and indepedent attribute(s) are not logically related, but model is showing otherwise.


```{r}
ts.plot(as.ts(data$Wisconsin_Employment_Rate), 
        as.ts(data$Milk_Production),
        ylab="Wisconsin Employent vs Milk production per cow")
```

```{r}
cor(data$Wisconsin_Employment_Rate, data$Milk_Production)
```

Certainly, there is correlation between the variables, but there is no causation.

The good model is because of **correlation** between variables and not because of **causation**.

# Spurious Regression
* Spurious Regression is a regression that provides misleading statistical evidence of a linear relationship between independent non-stationary variables
* Milk_Data and Emp_Data has a strong trend 
* Any other variable with such a trend will also show larger R^2^
* It shows apparent relationship between two variables but in reality it doesn't exist

```{r}
library(forecast)

ndiffs(data$Wisconsin_Employment_Rate)
ndiffs(data$Milk_Production)
```

```{r}
lr_diff = lm(diff(data$Wisconsin_Employment_Rate, lag=1)~diff(data$Milk_Production, lag=1))
summary(lr_diff)
```