---
title: "Predicting Labour Wages using Elastic Net Regression"
author: "Insofe Labs"
output:
  html_document:
    toc: yes
    toc_depth: '3'
---


# ElasticNet Regression

$$\text{Residual Mean Square Error}+\ (1-\alpha) \cdot \text{Ridge Penalty}+ \alpha\cdot \text{LASSO Penalty}$$

# Read and Understand the data

```{r}

labour_data <- read.csv("labour_income.csv")

str(labour_data)

summary(labour_data)

```

# Data Pre-processing

## Train-Test Split

* Split the data into train and test

```{r}

set.seed(007)

train_rows <- sample(x = seq(1, nrow(labour_data), 1), size = 0.7*nrow(labour_data))

train_data <- labour_data[train_rows, ]

test_data <- labour_data[-train_rows, ]

```

## Standardize the Data

* Standardize the continuous independent variables

```{r}

library(caret)

std_obj <- preProcess(x = train_data[, !colnames(train_data) %in% c("wages")],
                      method = c("center", "scale"))

train_std_data <- predict(std_obj, train_data)

test_std_data <- predict(std_obj, test_data)

```


## Dummify the Data

* Use the dummyVars() function from caret to convert sex and age into dummy variables

```{r}

dummy_obj <- dummyVars( ~ . , train_std_data)

train_dummy_data <- as.data.frame(predict(dummy_obj, train_std_data))

test_dummy_data <- as.data.frame(predict(dummy_obj, test_std_data))

```


# Hyper-parameter Tuning using CARET for Elastic Net

```{r}

library(glmnet)

sampling_strategy <- trainControl(method = "repeatedcv", number = 5, repeats = 3)

elastic_net_model <- train(wages ~ ., train_dummy_data,
                           method = "glmnet", 
                           trControl = sampling_strategy,
                           metric = "RMSE",
                           tuneGrid = 
                             expand.grid(.alpha = c(0.0500000, 0.1178571, 0.1857143, 0.2535714, 0.3214286, 0.3892857, 0.4571429, 0.5250000, 0.5928571, 0.6607143, 0.7285714, 0.7964286, 0.8642857, 0.9321429, 1.0000000), 
                                         #seq(.05, 1, length = 15),
                                         
                                  .lambda = c(0.1, 0.2, 0.3, 0.4, 0.5)
                                         #c((1:5)/10)
                                    )
                           )

```

```{r}

plot(elastic_net_model)

```


## Evaluating the Final Elastic Net Regression Model

* Use the model to predict on test data

```{r}

preds_elastic <- predict(elastic_net_model, test_dummy_data)

```


# Model Performance Evaluation

## Elastic Net Regression Model Metrics

```{r}

library(DMwR)

regr.eval(trues = test_dummy_data[, 1], preds = preds_elastic)

```

















